A VLT/FORS2 Narrowband Imaging Search for MgII Emission Around z$\sim$0.7 Galaxies

### 2017-12-26 Manuscript/vaught_2017.tex
Reorganized Files

\begin{abstract}
Galactic-scale outflows are thought to be the primary mechanism in the removal of cool gas in star-forming galaxies. Presently, the mass and energy of these flows remain poorly constrained. One way to better constrain these parameters is to measure the spatial extent of the outflow; however, measuring the spatial extent of such outflows via spectral methods has been traditionally very difficult due to the faintness of emission lines tracing outflowing material. We present VLT/FORS2 narrowband imaging of 5 star forming galaxies at redshift z=0.67-0.69 in the GOODS-S field as part of an effort to spatially resolve large-scale outflows traced by MgII emission. Previous spectra of this sample have already revealed winds traced by MgII absorption. At our sample redshift, the MgII emission lines fall exactly within the FORS2 HeII+47 and HeII/3000+48 interference filters. The total integration time of 10 hrs obtained in each filter permits the analysis of the flow surface brightness and extent on scales over which MgII is typically detected in absorption ( i.e, projected distances > 100 kpc). Such measurements can provide stronger constraints on the mass and energy of feedback, helping to advance our understanding of the processes regulating galaxy evolution up to z=1. 
\end{abstract}

@@

\section{Introduction}
Galactic scale outflows are considered to be the main driver of galaxy evolution, as they provide a means to transport cool gas away from star forming regions out to the intergalactic medium. In other words, this removal of cool gas carries away the necessary fuel that is required for stellar formation. Even though outflows play a critical role in the star formation rates and mass of a the host galaxy\citep{Werk_2014}, the physical production mechanism that powers these outflows remains uncertain. Despite the uncertainty, some possible mechanics have been proposed by theoretical studies that include thermal pressure from core collapse supernova, radiation pressure from starbursts and finally cosmic ray pressure\citep{Sugahara_2017,Larson_1974,Chevalier_1985,Springel_2003}. An accurate picture of what types of galaxies host outflows comes from numerous spectroscopic studies(authors), as outflows are detected by measuring the blueshift of absorption transitions with respect to the host galaxy systematic velocities. Spectroscopy of galaxies from small to large redshifts and over a wide range of star formation sequence have revealed outflows in galaxies which host a large concentrations of massive stars (e.g., \cite{Rubin_2014} Martin 2012).

One common practice in detecting the presence of gas in distant galaxies has been to spectroscopically probe galactic halos along background quasar sight lines and down-the-barrel observations. With temperatures of the cool halo gas component on the order of $10^4\ K$, the gas can be traced by the absorption of metals such \ion{Mg}{2}, Fe and others \citep{Bordoloi_2011, Bergeron_1986}.  Nonetheless, this method is weak in constraining the overall radial profile, extent and morphology of the outflow. It follows from these weak constraints that estimates of mass and energy loss are left uncertain by at least two order of magnitude \citep{Rubin_2011}.

@@

\section{Observations and Data Reduction}
Our sample consists of 5 GOODS-S galaxies. These galaxies were selected from a Keck/LRIS survey of UV absorption lines in 100 objects having redshifts $0.3< z , 1.4$ and $B_{AB}< 23$ in fields with deep HST/ACS imaging \cite{Rubin_2014}. The full sample is fortuitously distributed on the sky such that it can be imaged in a single FORS2 pointing.

The data were taken in service-mode using the FORS2 instrument on the VLT 8.2m telescope Antu between October 2012 and February 2013. We used two narrowband filters, HeII+47 and HeII/3000+48, that have peak transmission at wavelengths that correspond to the MgII doublet lines at our sample redshift of $z\approx 0.7$, (see Table \ref{tab:filters}). The FORS2 has a pixel scale of $0.25''$ pixel$^{-1}$ and a field of view of $7'\times7'$. Summing the individual exposure times for each image results in a combined exposure time of $10.0$ hours each for the HeII+47 and HeII/3000+48 filters. We carried out our observations under photometric and thin cloud conditions (program ID: 090.A-0427A).

\begin{figure}[!h]
\centering
\includegraphics[scale=.55]{sample_v2.png}
\caption{HST/ACS imaging of the each galaxy in our sample. From Top Left to Right Bottom: J033230.03, J033230.64, J033225.26, J033231.36, J033230.57.}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[scale=.55]{avg_seeing_HEII.pdf}
\includegraphics[scale=.55]{avg_seeing_HEII3000.pdf}
\caption{ The seeing conditions were calculated by ESO and were provided in the header of each science image. \textbf{(Top)} Distribution of seeing for the HeII+47 images. The median seeing is $1.03''$ \textbf{(Bottom)} : Same for the HEII/3000+48 images. The median seeing is $1.08''$ }
\end{figure}

The data were reduced using custom routines written in \textbf{Python}. The images were first corrected by subtracting and removing the overscan region of the CCD. Then the images were bias-subtracted and flat-fielded using twilight flats. Each individual image was cleaned of cosmic rays and bad pixels by utilizing the L.A Cosmic algorithm. The astrometry solutions were calculated via Astrometry.net \citep{Lang} with a standard deviation in the galaxy coordinates of $\sigma \approx 0.10''$.

Before image stacking, we ran each frame through \textbf{SExtractor} \citep{Bertin} in order to create a Root-Mean-Square map of each science image. These frames, in addition to the science images, are stacked to create a co-added RMS image.

The final stacked images for each filter is obtained using \textbf{SWarp} \citep{Bertin}. Each individual frame is first sky-subtracted using a background mesh size of 256 pixels which is approximately $64''$. We chose the mesh size to be large enough such that the extended emission is not mistakenly subtracted \citep{Battaia_2015}. The frame, after background-subtraction, is resampled onto a common astrometric solution using a \textit{Lancosz3} interpolation kernel. The images are median-combined in order to increase the signal to noise of any MgII emission. \textbf{SWarp} is also configured to produce RMS weight map from our \textbf{SExtractor} weighted images.

@@

\begin{figure*}[ht!]
\centering
\includegraphics[scale=.61]{HEII_final.png}
\includegraphics[scale=.61]{HEII3000_final.png}
\caption{Top: HeII+47 image after stacking. Bottom: HeII3000+48 image after stacking. }
\label{fig:stacked_image}
\end{figure*}
\section{Analysis}

@@

\begin{figure*}
\centering
\gridline{\fig{J033225_26.pdf}{0.5\textwidth}{(J033225.26)}
          \fig{J033230_03.pdf}{0.5\textwidth}{(J033230.03)}}
\gridline{\fig{J033230_64.pdf}{0.5\textwidth}{(J033230.64)}
          \fig{J033231_36.pdf}{0.5\textwidth}{(J033231.36)}}
\caption{Surface brightness profiles for galaxies that exhibit flux decrements from MgII absorption. Photometry was performed with circular apertures with increasing inner and outer radii. The error bars are determined from a SWarp RMS map .}
\label{fig:sb_profiles}
\end{figure*}

\begin{figure*}[!htb]
\centering
\includegraphics[scale=0.7]{stamps.pdf}
\caption{ $10'' \times 10''$ images stamps. Bottom: Continuum flux in ergs/s/cm$^2$/arcsec. Top: Continuum subtracted MgII flux, absorption can be seen in 3 of 5 galaxies. Each stamp is scaled with symmetric limits around 0.}
\label{fig:stamp_images}
\end{figure*}

\begin{figure*}[!htb]
\centering
\includegraphics[scale=0.58]{filt_57_spectra.pdf}
\includegraphics[scale=0.58]{filt_64_spectra.pdf}
\includegraphics[scale=0.58]{filt_03_spectra.pdf}
\includegraphics[scale=0.58]{filt_26_spectra.pdf}
\includegraphics[scale=0.58]{filt_36_spectra.pdf}
\caption{KECK/LRIS spectra plotted with the transmission curves of the filters HeII+47 and HeII/3000+48. The MgII doublet falls fortuitously the central wavelength of the HEII+47 filter. }
\label{fig:spec_images}
\end{figure*}

\begin{figure*}[ht!]
\centering
\includegraphics[scale=.58]{J03EW.pdf}
\includegraphics[scale=.58]{J64EW.pdf}
\includegraphics[scale=.58]{J26EW.pdf}
\includegraphics[scale=.58]{J57EW.pdf}
\includegraphics[scale=.58]{J36EW.pdf}
\caption{ Plots of the distribution of equivalent widths inside a $1\sigma$ continuum flux outline. }
\label{fig:ew_images}
\end{figure*}

>>>>

\begin{abstract}
Galactic-scale outflows are thought to be the primary mechanism in the removal of cool gas in star-forming galaxies. 
%
Presently, the mass and energy of these flows remain poorly constrained.
%
One way to better constrain these parameters is to measure the spatial extent of the outflow; however, measuring the spatial extent of such outflows via spectral methods has been traditionally very difficult due to the faintness of emission lines tracing outflowing material. 
%
We present VLT/FORS2 narrowband imaging of 5 star forming galaxies at redshift z=0.67-0.69 in the GOODS-S field as part of an effort to spatially resolve large-scale outflows traced by MgII emission. 
%
Previous spectra of this sample have already revealed winds traced by MgII absorption. 
%
At our sample redshift, the MgII emission lines fall exactly within the FORS2 HeII+47 and HeII/3000+48 interference filters.
%
The total integration time of 10 hrs obtained in each filter permits the analysis of the flow surface brightness and extent on scales over which MgII is typically detected in absorption ( i.e, projected distances > 100 kpc). 
%
Such measurements can provide stronger constraints on the mass and energy of feedback, helping to advance our understanding of the processes regulating galaxy evolution up to z=1. 
\end{abstract}

@@

\section{Introduction}
Galactic scale outflows are considered to be the main driver of galaxy evolution, as they provide a means to transport cool gas away from star forming regions out to the intergalactic medium. 
%
%
In other words, this removal of cool gas carries away the necessary fuel that is required for stellar formation. 
%
%
Even though outflows play a critical role in the star formation rates and mass of a the host galaxy\citep{Werk_2014}, the physical production mechanism that powers these outflows remains uncertain.
%
%
Despite the uncertainty, some possible mechanics have been proposed by theoretical studies that include thermal pressure from core collapse supernova, radiation pressure from starbursts and finally cosmic ray pressure\citep{Sugahara_2017,Larson_1974,Chevalier_1985,Springel_2003}. 
%
%
An accurate picture of what types of galaxies host outflows comes from numerous spectroscopic studies(authors), as outflows are detected by measuring the blueshift of absorption transitions with respect to the host galaxy systematic velocities.
%
%
Spectroscopy of galaxies from small to large redshifts and over a wide range of star formation sequence have revealed outflows in galaxies which host a large concentrations of massive stars (e.g., \cite{Rubin_2014} Martin 2012). 
%
% New Paragraph
One common practice in detecting the presence of gas in distant galaxies has been to spectroscopically probe galactic halos along background quasar sight lines and down-the-barrel observations.
%
With temperatures of the cool halo gas component on the order of $10^4\ K$, the gas can be traced by the absorption of metals such \ion{Mg}{2}, Fe and others \citep{Bordoloi_2011, Bergeron_1986}. 
%
Nonetheless, this method is weak in constraining the overall radial profile, extent and morphology of the outflow. 
%
It follows from these weak constraints that estimates of mass and energy loss are left uncertain by at least two order of magnitude \citep{Rubin_2011}.

@@

\section{Observations and Data Reduction}
Our sample consists of 5 GOODS-S galaxies. These galaxies were selected from a Keck/LRIS survey of UV absorption lines in 100 objects having redshifts $0.3< z , 1.4$ and $B_{AB}< 23$ in fields with deep HST/ACS imaging \cite{Rubin_2014}. 
%
The full sample is fortuitously distributed on the sky such that it can be imaged in a single FORS2 pointing.
% New Paragraph
The data were taken in service-mode using the FORS2 instrument on the VLT 8.2m telescope Antu between October 2012 and February 2013. 
%
We used two narrowband filters, HeII+47 and HeII/3000+48, that have peak transmission at wavelengths that correspond to the MgII doublet lines at our sample redshift of $z\approx 0.7$, (see Table \ref{tab:filters}).
%
The FORS2 has a pixel scale of $0.25''$ pixel$^{-1}$ and a field of view of $7'\times7'$.
%
Summing the individual exposure times for each image results in a combined exposure time of $10.0$ hours each for the HeII+47 and HeII/3000+48 filters.
%
We carried out our observations under photometric and thin cloud conditions (program ID: 090.A-0427A).

\begin{figure}[!h]
\centering
\includegraphics[scale=.55]{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/sample_v2.png}
\caption{HST/ACS imaging of the each galaxy in our sample. From Top Left to Right Bottom: J033230.03, J033230.64, J033225.26, J033231.36, J033230.57.}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[scale=.55]{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/avg_seeing_HEII.pdf}
\includegraphics[scale=.55]{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/avg_seeing_HEII3000.pdf}
\caption{ The seeing conditions were calculated by ESO and were provided in the header of each science image. \textbf{(Top)} Distribution of seeing for the HeII+47 images. The median seeing is $1.03''$ \textbf{(Bottom)} : Same for the HEII/3000+48 images. The median seeing is $1.08''$ }
\end{figure}

The data were reduced using custom routines written in \textbf{Python}. 
%%
The images were first corrected by subtracting and removing the overscan region of the CCD. 
%%
Then the images were bias-subtracted and flat-fielded using twilight flats. 
%%
Each individual image was cleaned of cosmic rays and bad pixels by utilizing the L.A Cosmic algorithm.
%%
 The astrometry solutions were calculated via Astrometry.net \citep{Lang} with a standard deviation in the galaxy coordinates of $\sigma \approx 0.10''$. 
%%
Before image stacking, we ran each frame through \textbf{SExtractor} \citep{Bertin} in order to create a Root-Mean-Square map of each science image.
%%
 These frames, in addition to the science images, are stacked to create a co-added RMS image.
%% New Paragraph
The final stacked images for each filter is obtained using \textbf{SWarp} \citep{Bertin}.
%%
Each individual frame is first sky-subtracted using a background mesh size of 256 pixels which is approximately $64''$. 
%%
We chose the mesh size to be large enough such that the extended emission is not mistakenly subtracted \citep{Battaia_2015}. 
%%
The frame, after background-subtraction, is resampled onto a common astrometric solution using a \textit{Lancosz3} interpolation kernel. 
%%
The images are median-combined in order to increase the signal to noise of any MgII emission. \textbf{SWarp} is also configured to produce RMS weight map from our \textbf{SExtractor} weighted images.

@@

\begin{figure*}[ht!]
\centering
\includegraphics[scale=.61]{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/HEII_final.png}
\includegraphics[scale=.61]{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/HEII3000_final.png}
\caption{Top: HeII+47 image after stacking. Bottom: HeII3000+48 image after stacking. }
\label{fig:stacked_image}
\end{figure*}
\section{Analysis}

@@

\begin{figure*}
\centering
\gridline{\fig{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/J033225_26.pdf}{0.5\textwidth}{(J033225.26)}
          \fig{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/J033230_03.pdf}{0.5\textwidth}{(J033230.03)}}
\gridline{\fig{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/J033230_64.pdf}{0.5\textwidth}{(J033230.64)}
          \fig{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/J033231_36.pdf}{0.5\textwidth}{(J033231.36)}}
\caption{Surface brightness profiles for galaxies that exhibit flux decrements from MgII absorption. Photometry was performed with circular apertures with increasing inner and outer radii. The error bars are determined from a SWarp RMS map .}
\label{fig:sb_profiles}
\end{figure*}

\begin{figure*}[!htb]
\centering
\includegraphics[scale=0.7]{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/stamps.pdf}
\caption{ $10'' \times 10''$ images stamps. Bottom: Continuum flux in ergs/s/cm$^2$/arcsec. Top: Continuum subtracted MgII flux, absorption can be seen in 3 of 5 galaxies. Each stamp is scaled with symmetric limits around 0.}
\label{fig:stamp_images}
\end{figure*}

\begin{figure*}[!htb]
\centering
\includegraphics[scale=0.58]{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/filt_57_spectra.pdf}
\includegraphics[scale=0.58]{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/filt_64_spectra.pdf}
\includegraphics[scale=0.58]{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/filt_03_spectra.pdf}
\includegraphics[scale=0.58]{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/filt_26_spectra.pdf}
\includegraphics[scale=0.58]{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/filt_36_spectra.pdf}
\caption{KECK/LRIS spectra plotted with the transmission curves of the filters HeII+47 and HeII/3000+48. The MgII doublet falls fortuitously the central wavelength of the HEII+47 filter. }
\label{fig:spec_images}
\end{figure*}

\begin{figure*}[ht!]
\centering
\includegraphics[scale=.58]{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/J03EW.pdf}
\includegraphics[scale=.58]{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/J64EW.pdf}
\includegraphics[scale=.58]{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/J26EW.pdf}
\includegraphics[scale=.58]{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/J57EW.pdf}
\includegraphics[scale=.58]{/Users/ryan/Desktop/papers/rrvaught_2018/Figures/J36EW.pdf}
\caption{ Plots of the distribution of equivalent widths inside a $1\sigma$ continuum flux outline. }
\label{fig:ew_images}
\end{figure*}
